Spin qudit tomography and reconstruction error

### 2021-07-29 sun_phases/spin_qudit_tomo.tex
minor updates to tomography paper

\title{Spin qudit tomography}
\author{Michael A.~Perlin}
\email{mika.perlin@gmail.com}
\author{Diego Barberena}
\author{Ana Maria Rey}
\affiliation{JILA, National Institute of Standards and Technology and University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
\affiliation{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}

@@

Due to advancements in experimental capabilities to address nuclear spin states (i.e.~hyperfine levels) in ultracold atomic systems \cite{daley2011quantum, mischuck2012control, smith2013quantum, cazalilla2014ultracold}, as well as developments in the control of ultracold molecular systems \cite{bohn2017cold, zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{hemmer2018squeezing, evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, % todo: cite SU(n) spin model paper!
and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2021designing}.
In all cases, quantum state tomography is necessary to take full advantage of a spin qudit.

% todo: explain the role of tomography in sensing and simulation?

The problem of qudit tomography is not new, with an extensive literature on a variety of techniques \cite{newton1968measurability, hofmann2004quantumstate, filippov2010inverse, schmied2011tomographic, evrard2019enhanced, flammia2005minimal, thew2002qudit, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam, filippov2010inverse, hofmann2004quantumstate, filippov2010inverse}.
However, most existing protocols either rely on infinite-dimensional representations of a quantum spin \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced}, or require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, generally resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
The protocols based on infinite-dimensional representations of a quantum spin have the advantage of reconstructing its state from measurements of spin projection onto different spatial axes, which are generally accessible with any spin qudit.
Nonetheless, in addition to their inefficiencies these protocols obfuscate the minimal requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and (with the notable exception of Ref.~\cite{schmied2011tomographic}) generally extract only a small fraction of the information contained in measurement data.

In this work, we consider the task of performing spin qubit tomography using only measurements of spin projection onto different spatial axes, and address some of the shortcomings discussed above.
This sort of qudit tomography was first considered in Ref.~\cite{newton1968measurability}, as well as a few later works \cite{hofmann2004quantumstate, filippov2010inverse, schmied2011tomographic}.
Specifically, Ref.~\cite{newton1968measurability} provided an explicit protocol for reconstructing a $d$-level spin qudit state from measurements of spin projection along $2d-1$ axes, the minimum number necessary for full tomography of an arbitrary qudit state.
However, the protocol in Ref.~\cite{newton1968measurability} involves a choice of a single (arbitrary) angle $\theta$, and provides no means for comparing different choices of $\theta$, which may result in wildly different statistical errors (i.e.~precision) in a reconstructed state.
Other works provide insightful discussions into the problem of spin qudit tomography, but either
\begin{enumerate*}
\item require making assumptions about the qudit state in question \cite{schmied2011tomographic},
\item do not address the question of statistical error \cite{hofmann2004quantumstate}, or
\item provide a measure of statistical error that is overly conservative and computationally demanding \cite{filippov2010inverse}.
\end{enumerate*}

In Section \ref{sec:polarization_ops}, we introduce a set of qudit operators that are closely related to the spherical harmonics, and which play a central role in our work.
We then map the quantum problem of spin qudit tomography onto the classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby providing an interesting new perspective on spin qudit tomography.
In Section \ref{sec:error} we provide a priori upper bounds and a posteriori estimates of the statistical error in a qudit state reconstructed from measurements of spin projection along a given set of $r$ measurement axes.
The capability to determine upper bounds on reconstruction error a priori motivates a simple randomized tomography protocol that we outline in Section \ref{sec:protocol}, and for which we numerically find that using more measurement axes yields substantial benefits that plateau after $r\approx3d$.
To facilitate the use of our protocols, we make all of our codes publicly available at Ref.~\cite{tomo_codes}, which also contains the best measurement axes we found for $d\le20$ and $r=3d$.

@@

Here $\bm v$ and $\ell$ label a row of $M_V$, or equivalently label a standard basis vector $\ket{\bm v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space, and $\obra{T_{\bm v\ell,0}}$ is the conjugate transpose of $\oket{T_{\bm v\ell,0}}$.
A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators in $\set{T_{\bm v\ell,0}}$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.

@@

where, using the fact that $T_j=T_j^\dag$ (because they are degree $m=0$ polarization operators),

@@

Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$:

@@

  \bbk{\epsilon_j \epsilon_{j'}} \le \epsilon^2 \delta_{jj'},

@@

where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging over experimental trials that estimate $\bk{T_j}_\rho$.

@@

where $\norm{X}^2 \equiv \obk{X|X} = \tr\p{X^\dag X}$ is the squared (Frobenius, or Hilbert-Schmidt) norm of $X$, and the {\it classical error scale} $\S_V$ is defined by

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following protocol for spin qudit tomography:
\begin{enumerate}
\item Select a random set of measurement axes $V$ by uniformly sampling points on the sphere\footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and use any standard minimization algorithm to optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\bm v\in V$) by minimizing the quantum error scale $\epsilon_V$ in Eq.~\eqref{eq:bound}.
If $\abs{V}$ is too large for such optimization, you can simply generate many sets of random measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
Note that computing the error scale $\epsilon_V$ requires, for each $\ell\in\set{0,1,\cdots,d-1}$, constructing the measurement matrix $M_{V\ell}$ in Eq.~\eqref{eq:block} and computing its singular value decomposition.
Save all measurement matrix data associated with the final measurement axes $V$ for later use.
\item For each axis $\bm v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\bm v\mu}\approx\bk{\Pi_{\bm v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
\item Use the the estimates $\tilde\Pi_{\bm v\mu}$ of $\bk{\Pi_{\bm v\mu}}_\rho$ to compute the estimates
\begin{align}
  \tilde T_{\bm v\ell,0}
  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\bm v\mu}
\end{align}
of $\bk{T_{\bm v\ell,0}}_\rho$, where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
\item Denoting the nonzero singular values of $M_{V\ell}$ by $M^V_{\ell k}$ and the corresponding left singular vectors by $\bm x^V_{\ell k} = \sum_{\bm v} x^V_{\ell k \bm v} \ket{\bm v}$, compute the  operators and coefficients
\begin{align}
  Q_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\bm v}
  \p{x^V_{\ell k\bm v}}^* T_{\bm v\ell,0}, \\
  \tilde\rho^V_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\bm v}
  x^V_{\ell k\bm v} \, \tilde T_{\bm v\ell,0},
\end{align}
and combine them into the estimate
\begin{align}
  \tilde\rho_V = \sum_{\ell,k} \tilde\rho^V_{\ell k} Q_{\ell k}
  \approx \rho.
\end{align}
\end{enumerate}
The expected reconstruction error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is provided by Eq.~\eqref{eq:error_cls}.
If measurement error is shot-noise-limited, then the error in $\tilde\rho_V$ is approximately $\E_V\p{\rho}\approx\E_V\p{\tilde\rho_V}$ and can be computed from Eq.~\eqref{eq:error_qnt}.
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

In principle, any information about $\rho$ obtained from prior knowledge or preliminary measurement data can be used to construct tailored or adaptive measurement protocols \cite{huszar2012adaptive, ferrie2014selfguided, granade2016practical, pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
We leave the development of tailored and adaptive measurement protocols to future work.

@@

\begin{figure}
  \centering
  \includegraphics{qudit_axes.pdf}
  \caption{Empirical measurement-adjusted error scales $\tilde\beta(p)$ with $p$ excess measurement axes, determined by minimizing over $10^3$ choices of measurement axes $V$ or 5 minutes of runtime (for each $p$), whichever comes first.
    Color indicates the qudit dimension $d$.
    The rapid initial drop in $\tilde\beta(p)$ implies that using more measurement axes can substantially lower the upper bound on reconstruction error provided in Eq.~\eqref{eq:bound}, and that these benefits plateau after $p\approx d$.
    Horizontal reference lines mark the smallest measurement-adjusted error scales $\min_\theta\beta_\theta/\tilde\beta(0)$ achievable with the method in Ref.~\cite{newton1968measurability}, which is parameterized by an angle $\theta$.}
  \label{fig:axes}
\end{figure}

The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize the (squared) reconstruction error $\E_V\p{\rho}^2<\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
We therefore consider the measurement-adjusted error scale $\beta(p)$ defined by
\begin{align}
  \beta(p)^2
  \equiv \min_V\Set{\epsilon_V^2\abs{V}:\abs{V}=2d-1+p}
\end{align}
as a function of $p$, the number of measurement axes exceeding $2d-1$.
Though we cannot minimize over all suitable choices of measurement axes $V$ to compute $\beta(p)$, we can compute an empirical upper bound $\tilde\beta(p)\ge\beta(p)$ by minimizing over a large number of randomly chosen $V$.
Figure \ref{fig:axes} shows the results of such empirical minimization, where we find that $\tilde\beta(p)$ drops substantially with $p$ before plateauing at $p\approx d$, after which there are only minor benefits to using more measurement axes.
In the interest of reducing experimental complexity as well as the runtime of our randomized tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that this protocol should be performed with $\abs{V}\approx3d$ measurement axes.
We provide the best measurement axes that we found for a randomized tomography protocol with $d\le20$ and $\abs{V}=3d$ in Ref.~\cite{tomo_codes}.

For reference, Figure \ref{fig:axes} also shows the smallest measurement-adjusted error scales $\beta_\theta$ achievable with the method of Ref.~\cite{newton1968measurability}, which is comparable to those achieved with our randomized protocol at $\abs{V}\approx3d$ ($p\approx d$).
The method of Ref.~\cite{newton1968measurability} requires choosing an angle $\theta$, and provides no prescription for making this choice.
We therefore find the optimal choice of $\theta$ by minimizing $\beta_\theta$ over all $\theta$ (see Appendix \ref{sec:angle_search}), and show $\min\beta_\theta/\tilde\beta(0)$ in Figure \ref{fig:axes}.
Empirically, we find that the optimal angle for the method of Ref.~\cite{newton1968measurability} is $\theta_{\t{opt}}\approx\frac{\pi}{2}(1 - \frac1{1.34d})$ (see Appendix \ref{sec:angle_search}), although we note that a randomized tomography protocol is likely more robust to errors in axis orientation because measurement axes are generally more spaced out.
We leave the question of robustness to errors in axis orientation and the effect of these errors on state reconstruction error to future work.

@@

Ref.~\cite{newton1968measurability} constructs an explicit protocol for spin qudit tomography, which involves measuring spin projection along $2d-1$ axes equally spaced at a polar angle $\theta$.
However, this method does not provide any prescription for choosing $\theta$.
Here, we show the importance of making a good choice of $\theta$, and empirically find the optimal value of $\theta_{\t{opt}}$ that minimizes the corresponding quantum error scale $\epsilon_\theta$, which controls state reconstruction error.
To this end, Figure \ref{fig:angle_sweep} shows the quantum error scale $\epsilon_\theta$ as a function of the polar angle $\theta$ in the tomography method of Ref.~\cite{newton1968measurability} for a few qudit dimensions $d$.
While a good choice of $\theta$ yields an error scale $\epsilon_\theta\approx d$ (for the dimensions shown), this error scale can increase by orders of magnitude for poor choices of $\theta$.
In turn, Figure \ref{fig:opt_angles} shows the optimal angle $\theta_{\t{opt}}$ as a function of the qudit dimension $d$, together with a fit to $\theta_{\t{opt}}=\frac{\pi}{2}(1-\frac1{xd})$ finding $x\approx1.34\pm0.008$.

@@

\begin{figure}
  \centering
  \includegraphics{opt_angles.pdf}
  \caption{Optimum angle $\theta_{\t{opt}}$ as a function of qudit dimension $d$ for the tomography method of Ref.~\cite{newton1968measurability}, and a fit to $\theta_{\t{opt}}=\frac{\pi}{2}(1-\frac1{xd})$ finding $x\approx1.34\pm0.008$.}
  \label{fig:opt_angles}
\end{figure}

>>>>

\title{Spin qudit tomography and reconstruction error}
\author{Michael A.~Perlin}
\email{mika.perlin@gmail.com}
\author{Diego Barberena}
\author{Ana Maria Rey}
\affiliation{JILA, National Institute of Standards and Technology and University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
\affiliation{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}

@@

Due to advancements in experimental capabilities to address nuclear spin states (i.e.~hyperfine levels) in ultracold atomic systems \cite{daley2011quantum, mischuck2012control, smith2013quantum, cazalilla2014ultracold}, as well as developments in the control of ultracold molecular systems \cite{bohn2017cold, zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.  Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{hemmer2018squeezing, evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, % todo: cite SU(n) spin model paper!
and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2021designing}.
In all cases, quantum state tomography is necessary to take full advantage of a spin qudit\footnote{Note that the measurement of collective observables for quantum sensing or simulation can be recast as a single-spin tomography task.}.

The problem of qudit tomography is not new, with an extensive literature on a variety of techniques \cite{newton1968measurability, hofmann2004quantumstate, filippov2010inverse, schmied2011tomographic, evrard2019enhanced, flammia2005minimal, thew2002qudit, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam, filippov2010inverse, hofmann2004quantumstate, filippov2010inverse}.
However, most existing protocols either rely on infinite-dimensional representations of a quantum spin \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced}, or require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, generally resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
The protocols based on infinite-dimensional representations of a quantum spin have the advantage of reconstructing its state from measurements of spin projection onto different spatial axes, which are generally accessible with any spin qudit.
Nonetheless, these protocols obfuscate the minimal requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and (with the notable exception of Ref.~\cite{schmied2011tomographic}) generally extract only a small fraction of the information contained in measurement data.

In this work, we consider the task of performing spin qubit tomography using only measurements of spin projection onto different spatial axes.
This sort of task was first considered in Ref.~\cite{newton1968measurability}, as well as a few later works \cite{hofmann2004quantumstate, filippov2010inverse, schmied2011tomographic}.
Specifically, Ref.~\cite{newton1968measurability} provided an explicit protocol for reconstructing a $d$-level spin qudit state from measurements of spin projection along $2d-1$ axes, the minimum number necessary for full tomography of an arbitrary (possibly mixed) qudit state.
However, the protocol in Ref.~\cite{newton1968measurability} involves a choice of a single (arbitrary) angle $\theta$, and provides no means for comparing different choices of $\theta$, which may result in wildly different statistical errors (i.e.~precision) in a reconstructed state.
Other works provide insightful discussions into the problem of spin qudit tomography, but either
\begin{enumerate*}
\item require making assumptions about the qudit state in question \cite{schmied2011tomographic} (making the tomographic protocol only valid for a restricted set of possible states),
\item do not address the question of statistical error \cite{hofmann2004quantumstate}, or
\item provide a measure of statistical error that is needlessly conservative and computationally demanding \cite{filippov2010inverse}.
\end{enumerate*}
We address these shortcomings in this work, and identify remaining avenues for refining spin qudit tomography protocols.

In Section \ref{sec:polarization_ops}, we introduce a set of qudit operators that are closely related to the spherical harmonics, and which play a central role in our work.
We then map the quantum problem of spin qudit tomography onto the classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby providing an interesting new perspective on spin qudit tomography.
In Section \ref{sec:error} we provide a priori upper bounds and a posteriori estimates of the statistical error in a qudit state reconstructed from measurements of spin projection along a given set of $r$ measurement axes.
The capability to determine upper bounds on reconstruction error a priori motivates a simple randomized tomography protocol that we outline in Section \ref{sec:protocol}, and for which we numerically find that using more measurement axes yields substantial benefits that plateau after $r\approx3d$.
To facilitate the use of our protocols, we make all of our codes publicly available at Ref.~\cite{tomo_codes}, which also contains the best measurement axes we found for $d\le30$ and $r=3d$.

@@

Here $\bm v$ and $\ell$ label a row of $M_V$, or equivalently label a standard (``one-hot'') basis vector $\ket{\bm v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space, and $\obra{T_{\bm v\ell,0}}$ is the conjugate transpose of $\oket{T_{\bm v\ell,0}}$.
A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators $T_{\bm v\ell,0}$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.

@@

where, using the fact that $T_j=T_j^\dag$ (because they are diagonal, degree $m=0$ polarization operators),

@@

Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance no grater than $\epsilon^2$:

@@

  \bbk{\epsilon_j \epsilon_{j'}} \le \epsilon^2 \delta_{jj'}.

@@

Here $\set{\epsilon_j}$ are independent random variables, and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging over experimental trials that estimate $\bk{T_j}_\rho$.

@@

where $\norm{X}^2 \equiv \obk{X|X} = \tr\p{X^\dag X}$ is the squared (Euclidean, Frobenius, or Hilbert-Schmidt) norm of $X$, and the {\it classical error scale} $\S_V$ is defined by

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following protocol for spin qudit tomography:
\begin{enumerate}
\item Select a random set of measurement axes $V$ by uniformly sampling points on the sphere\footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and use any standard minimization algorithm to optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\bm v\in V$) by minimizing the quantum error scale $\epsilon_V$ in Eq.~\eqref{eq:bound}.
If $\abs{V}$ is too large for such optimization, you can simply generate many sets of random measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
Note that computing the error scale $\epsilon_V$ requires, for each $\ell\in\set{0,1,\cdots,d-1}$, constructing the measurement matrix $M_{V\ell}$ in Eq.~\eqref{eq:block} and computing its singular value decomposition.
Save all measurement matrix data associated with the final measurement axes $V$ for later use.
\item For each axis $\bm v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\bm v\mu}\approx\bk{\Pi_{\bm v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
\item Use the the estimates $\tilde\Pi_{\bm v\mu}$ of $\bk{\Pi_{\bm v\mu}}_\rho$ to compute the estimates
\begin{align}
  \tilde T_{\bm v\ell,0}
  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\bm v\mu}
\end{align}
of $\bk{T_{\bm v\ell,0}}_\rho$, where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
\item Denoting the nonzero singular values of $M_{V\ell}$ by $M^V_{\ell k}$ and the corresponding left singular vectors by $\bm x^V_{\ell k} = \sum_{\bm v} x^V_{\ell k \bm v} \ket{\bm v}$, compute the  operators and coefficients
\begin{align}
  Q_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\bm v}
  \p{x^V_{\ell k\bm v}}^* T_{\bm v\ell,0}, \\
  \tilde\rho^V_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\bm v}
  x^V_{\ell k\bm v} \, \tilde T_{\bm v\ell,0},
\end{align}
and combine them into the estimate
\begin{align}
  \tilde\rho_V = \sum_{\ell,k} \tilde\rho^V_{\ell k} Q_{\ell k}
  \approx \rho.
\end{align}
\end{enumerate}
The expected reconstruction error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is provided by Eq.~\eqref{eq:error_cls}.
If measurement error is shot-noise-limited, then the error in $\tilde\rho_V$ is approximately $\E_V\p{\rho}\approx\E_V\p{\tilde\rho_V}$ and can be computed from Eq.~\eqref{eq:error_qnt}.
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}, which will additionally guarantee that $\tilde\rho_V$ satisfies all requirements for being a physical state.

@@

\begin{figure}
  \centering
  \includegraphics{qudit_axes.pdf}
  \caption{Empirical measurement-adjusted error scales $\tilde\beta(p)$ with $p$ excess measurement axes, determined by minimizing over $10^3$ choices of measurement axes $V$ or 5 minutes of runtime (for each $p$), whichever comes first.
    Color indicates the qudit dimension $d$.
    The rapid initial drop in $\tilde\beta(p)$ implies that using more measurement axes can substantially lower the upper bound on reconstruction error provided in Eq.~\eqref{eq:bound}, and that these benefits plateau after $p\approx d$.
    Horizontal reference lines mark the smallest measurement-adjusted error scales $\min_\theta\beta_\theta/\tilde\beta(0)$ achievable with the method in Ref.~\cite{newton1968measurability}, which is parameterized by an arbitrarily chosen angle $\theta$.}
  \label{fig:axes}
\end{figure}

The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize the (squared) reconstruction error $\E_V\p{\rho}^2<\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
We therefore consider the measurement-adjusted error scale $\beta(p)$ defined by
\begin{align}
  \beta(p)^2
  \equiv \min_V\Set{\epsilon_V^2\abs{V}:\abs{V}=2d-1+p},
\end{align}
where $p$ is the number of ``extra'' measurement axes exceeding $2d-1$.
Though we cannot minimize over all suitable choices of measurement axes $V$ to compute $\beta(p)$, we can compute an empirical upper bound $\tilde\beta(p)\ge\beta(p)$ by minimizing over a large number of randomly chosen $V$.
Figure \ref{fig:axes} shows the results of such empirical minimization, where we find that $\tilde\beta(p)$ drops substantially with $p$ before plateauing at $p\approx d$, after which there are only minor benefits to using more measurement axes.
In the interest of reducing experimental complexity as well as the runtime of our randomized tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that this protocol should be performed with $\abs{V}\approx3d$ measurement axes.
We provide the best measurement axes that we found for a randomized tomography protocol with $d\le30$ and $\abs{V}=3d$ in Ref.~\cite{tomo_codes}.

For reference, Figure \ref{fig:axes} also shows the smallest measurement-adjusted error scales $\beta_\theta$ achievable with the method of Ref.~\cite{newton1968measurability}, which is comparable to those achieved with our randomized protocol at $\abs{V}\approx3d$.
The method of Ref.~\cite{newton1968measurability} requires choosing an angle $\theta$, and provides no prescription for making this choice.
We therefore find the optimal choice of $\theta$ by minimizing the error scale $\beta_\theta$ over all $\theta$ (see Appendix \ref{sec:angle_search}), and show $\min\beta_\theta/\tilde\beta(0)$ in Figure \ref{fig:axes}.
Empirically, we find that the optimal angle for the method of Ref.~\cite{newton1968measurability} is $\theta_{\t{opt}}\approx\frac{\pi}{2}(1 - \frac1{1.34d})$ (see Appendix \ref{sec:angle_search}), although we note that a randomized tomography protocol is likely more robust to errors in axis orientation because measurement axes are generally more spaced out.
We leave the question of robustness to errors in axis orientation and the effect of these errors on state reconstruction error to future work.

As a final point, we note that any information about an unknown qudit state $\rho$, obtained from prior knowledge or preliminary measurement data, can be used to construct tailored or adaptive measurement protocols \cite{huszar2012adaptive, ferrie2014selfguided, granade2016practical, pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
We leave the development of tailored and adaptive measurement protocols to future work.

@@

Ref.~\cite{newton1968measurability} constructs an explicit protocol for spin qudit tomography, which involves measuring spin projection along $2d-1$ axes equally spaced at a polar angle $\theta$.
However, this method does not provide any prescription for choosing $\theta$.
Here, we show the importance of making a good choice of $\theta$, and empirically find the optimal value of $\theta_{\t{opt}}$ that minimizes the corresponding quantum error scale $\epsilon_\theta$, which controls state reconstruction error.
To this end, Figure \ref{fig:angle_sweep} shows the quantum error scale $\epsilon_\theta$ as a function of the polar angle $\theta$ in the tomography method of Ref.~\cite{newton1968measurability} for a few qudit dimensions $d$.
While a good choice of $\theta$ yields an error scale $\epsilon_\theta\approx d$ (for the dimensions shown), this error scale can increase by orders of magnitude for poor choices of $\theta$.
In turn, Figure \ref{fig:opt_angles} shows the optimal angle $\theta_{\t{opt}}$ as a function of the qudit dimension $d$, together with a fit to $\theta_{\t{opt}}=\frac{\pi}{2}(1-\frac1{xd})$ finding $x\approx1.34$.

@@

\begin{figure}
  \centering
  \includegraphics{opt_angles.pdf}
  \caption{Optimum angle $\theta_{\t{opt}}$ as a function of qudit dimension $d$ for the tomography method of Ref.~\cite{newton1968measurability}, and a fit to $\theta_{\t{opt}}=\frac{\pi}{2}(1-\frac1{xd})$ finding $x\approx1.34$.}
  \label{fig:opt_angles}
\end{figure}
